Event History Analyses for psychological time-to-event data: A tutorial in R with examples in Bayesian and frequentist workflows

### 2024-10-25 manuscript/manuscript.Rmd
update ms. and extrareferences2.bib

Descriptive, data science workflow.
Data wrangling via tidyverse principles and a functional programming approach (cite R4DS textbook here).
Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
[[more here, as necessary]].

@@

Our lab adopts a estimation approach to multi-level regression (Kruschke & Liddel, 2018; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2020; Kurz, 202?). We also use a "keep it maximal" approach to specifying varying (or random) effects (Barr et al., 2013). This means that wherever possible we include varying intercepts and slopes per participant 
To make inferences, we use two main approaches. We compare models of different complexity, using information criteria, such as WAIC or LOO, to evaluate out-of-sample predictive accuracy. We also take the most complex model and evaluate key parameters of interest using point and interval estimates.

>>>>

Descriptive, data science workflow.
Data wrangling via tidyverse principles and a functional programming approach [@wickhamDataScienceImport2023].
Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
[[more here, as necessary]].

@@

Our lab adopts a estimation approach to multi-level regression [@kruschkeBayesianNewStatistics2018] ; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath [@mcelreathStatisticalRethinkingBayesian2018] ; Kurz, 202?). We also use a "keep it maximal" approach to specifying varying (or random) effects [@barrRandomEffectsStructure2013]. This means that wherever possible we include varying intercepts and slopes per participant 
To make inferences, we use two main approaches. We compare models of different complexity, using information criteria, such as WAIC or LOO, to evaluate out-of-sample predictive accuracy. We also take the most complex model and evaluate key parameters of interest using point and interval estimates.

### 2024-10-25 manuscript/manuscript.tex
update ms. and extrareferences2.bib

Descriptive, data science workflow.
Data wrangling via tidyverse principles and a functional programming approach (cite R4DS textbook here).
Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
{[}{[}more here, as necessary{]}{]}.

@@

Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddel, 2018; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2020; Kurz, 202?). We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr et al., 2013). This means that wherever possible we include varying intercepts and slopes per participant
To make inferences, we use two main approaches. We compare models of different complexity, using information criteria, such as WAIC or LOO, to evaluate out-of-sample predictive accuracy. We also take the most complex model and evaluate key parameters of interest using point and interval estimates.

@@

@@

@@

@@

\begin{itemize}
\tightlist
\item
  censor(df,timeout,bin\_width) : divide the time segment (0,timeout{]} in bins, identify any right-censored observations, and determine the discrete RT (time bin rank)
\item
  ptb(df) : transform the person-trial data set to the person-trial-bin data set
\item
  setup\_lt(ptb) : set up a life table for each level of 1 independent variable
\item
  setup\_lt\_2IV(ptb) : set up a life table for each combination of levels of 2 independent variables
\item
  calc\_ca(df) : estimate the conditinal accuracies when there is 1 independent variable
\item
  calc\_ca\_2IV(df) : estimate the conditional accuraies when there are 2 independent variables
\item
  join\_lt\_ca(df1,df2) : add the ca(t) estimates to the life tables (1 independent variable)
\item
  join\_lt\_ca\_2IV(df1,df2) : add the ca(t) estimates to the life tables (2 independent variables)
\item
  extract\_median(df) : estimate quantiles S(t).50 (1 independent variable)
\item
  extract\_median\_2IV(df) : estimate quantiles S(t).50 (2 independent variables)
\item
  plot\_eha(df,subj,haz\_yaxis) : create plots of the discrete-time functions (1 independent variable)
\item
  plot\_eha\_2IV(df,subj,haz\_yaxis) : create plots of the discrete-time functions (2 independent variables)
\item
  plot\_eha\_agg(df,subj,haz\_yaxis) : create 1 plot for aggregated data (1 independent variable)
\end{itemize}

>>>>

Descriptive, data science workflow.
Data wrangling via tidyverse principles and a functional programming approach (Wickham, Çetinkaya-Rundel, \& Grolemund, 2023).
Functional programming basically means you don't write your own loops but instead use functions that have been built and tested by others.
{[}{[}more here, as necessary{]}{]}.

@@

Our lab adopts a estimation approach to multi-level regression (Kruschke \& Liddell, 2018) ; Winter, 2019), which is heavily influenced by Bayesian approach as suggested by Richard McElreath (McElreath, 2018) ; Kurz, 202?). We also use a ``keep it maximal'' approach to specifying varying (or random) effects (Barr, Levy, Scheepers, \& Tily, 2013). This means that wherever possible we include varying intercepts and slopes per participant
To make inferences, we use two main approaches. We compare models of different complexity, using information criteria, such as WAIC or LOO, to evaluate out-of-sample predictive accuracy. We also take the most complex model and evaluate key parameters of interest using point and interval estimates.

@@

\bibitem[\citeproctext]{ref-barrRandomEffectsStructure2013}
Barr, D. J., Levy, R., Scheepers, C., \& Tily, H. J. (2013). Random effects structure for confirmatory hypothesis testing: {Keep} it maximal. \emph{Journal of Memory and Language}, \emph{68}(3), 10.1016/j.jml.2012.11.001. \url{https://doi.org/10.1016/j.jml.2012.11.001}

@@

\bibitem[\citeproctext]{ref-kruschkeBayesianNewStatistics2018}
Kruschke, J. K., \& Liddell, T. M. (2018). The {Bayesian New Statistics}: {Hypothesis} testing, estimation, meta-analysis, and power analysis from a {Bayesian} perspective. \emph{Psychonomic Bulletin \& Review}, \emph{25}(1), 178--206. \url{https://doi.org/10.3758/s13423-016-1221-4}

@@

\bibitem[\citeproctext]{ref-wickhamDataScienceImport2023}
Wickham, H., Çetinkaya-Rundel, M., \& Grolemund, G. (2023). \emph{R for data science: Import, tidy, transform, visualize, and model data} (2nd edition). Beijing Boston Farnham Sebastopol Tokyo: O'Reilly.

@@

\begin{itemize}
\tightlist
\item
  censor(df,timeout,bin\_width) : divide the time segment (0,timeout{]} in bins, identify any right-censored observations, and determine the discrete RT (time bin rank)
\item
  ptb(df) : transform the person-trial data set to the person-trial-bin data set
\item
  setup\_lt(ptb) : set up a life table for each level of 1 independent variable
\item
  setup\_lt\_2IV(ptb) : set up a life table for each combination of levels of 2 independent variables
\item
  calc\_ca(df) : estimate the conditinal accuracies when there is 1 independent variable
\item
  calc\_ca\_2IV(df) : estimate the conditional accuraies when there are 2 independent variables
\item
  join\_lt\_ca(df1,df2) : add the ca(t) estimates to the life tables (1 independent variable)
\item
  join\_lt\_ca\_2IV(df1,df2) : add the ca(t) estimates to the life tables (2 independent variables)
\item
  extract\_median(df) : estimate quantiles S(t)\textsubscript{.50} (1 independent variable)
\item
  extract\_median\_2IV(df) : estimate quantiles S(t)\textsubscript{.50} (2 independent variables)
\item
  plot\_eha(df,subj,haz\_yaxis) : create plots of the discrete-time functions (1 independent variable)
\item
  plot\_eha\_2IV(df,subj,haz\_yaxis) : create plots of the discrete-time functions (2 independent variables)
\item
  plot\_eha\_agg(df,subj,haz\_yaxis) : create 1 plot for aggregated data (1 independent variable)
\end{itemize}
